Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-12 rainbowCAM.tex
Finished the first draft of the Introduction.

Time-independent scattering by a compact body of radius $R$ with a regular centre, such as a neutron star or white dwarf, has received less attention, by comparison. In such a scenario, an electromagnetic wave will not penetrate far inside the compact body; but on the other hand, a gravitational wave will pass through the body without much impediment from the matter distribution. Similarly, a neutron star is expected to be substantially transparent to neutrinos. Even in cases where the coupling to matter is negligible, the incident wave will nevertheless be scattered by the influence of the spacetime curvature. Thus, the resulting scattering pattern depends not just on $M\omega$, but also on the internal structure of the body, and on its inverse compactness or \emph{tenuity}, $R/M$. Typical values of the tenuity are: $R/M \sim 6$ for neutron stars, $\sim 1.4-9.4 \times 10^3$ for white dwarfs, $4.7 \times 10^5$ for the Sun, and $1.4 \times 10^9$ for Earth. More speculatively, we shall also consider here the case of (hypothetical) \emph{ultra-compact objects} (UCOs), whose tenuity is bounded from below by the Buchdahl bound of $R / M = 9/4$.

Studies of black hole scattering typically involve the calculation of scattering amplitude(s) via partial-wave expansions.  A powerful approach for resumming these partial-wave expansions is provided by so-called Complex Angular Momentum (CAM) techniques. In the CAM approach, the sum over partial waves is replaced by an integral, and the contour of integration is deformed into  the complex-angular-momentum plane such that one collects a sum over the residues of simple poles: these are the so-called \emph{Regge poles}.

@@

CAM techniques are yet to be applied to compact-body scattering, it appears, with one notable exception: in 1991, Chandrasekhar and Ferrari \cite{ChandrasekharIV:1992ey} examined the Regge pole spectra for relativistic stellar models. 
Their study focussed on resonant modes (poles with small imaginary parts corresponding to trapped $w$-modes or fluid modes), and it employed CAM theory to calculate the flow of gravitational energy through a star. It seems reasonable to expect that CAM can aid our understanding of other compact objects such as and neutron stars and UCOs, and this is our purpose below.

There is a close relationship between Regge poles and quasi-normal modes (QNMs); both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in time-dependent scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a `ringdown phase' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well-studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

It was recently shown \cite{Dolan:2017rtj, Stratton:2019deq} that time-independent scattering by compact bodies generates (in principle) a \emph{rainbow scattering} phenomenon in the short wavelength regime ($M \omega \gg 1$). The rainbow scattering phenomenon is linked to the formation of a cusp caustic in the incident wavefront (see Fig.~1 in Ref.~\cite{Stratton:2019deq}), and to the stationary point in the geodesic deflection function associated with a ray (that is, a null geodesic) that passes somewhat inside the body. For dilute bodies, such as stars and white dwarfs, the cusp caustic forms at a distance $d \sim R^2 / (4M)$ downstream of the body, and the rainbow angle is $\theta_r \approx 4M/R$ (e.g.~for the Sun, $d \approx 550 \, \text{a.u.}$ and $\theta_r \approx 1.8 \, \text{arcsec}$) \footnote{The rainbow angle $\theta_r \approx 4M/R$ is distinct from the Einstein ring angle of $\theta_E \approx \sqrt{4M/r}$}. For compact bodies such as neutron stars, however, the cusp caustic forms near the surface of the body, and the rainbow angle is large ($\theta_r \gtrsim 59.6^\circ$ for $R/M = 6$; see Table I in Ref.~\cite{Stratton:2019deq}). In this case, the rainbow angle is -- in principle at least -- a useful diagnostic of the matter distribution within the body and thus its nuclear equation of state.

@@

Finally a paragraph on the connections to Mie scattering by a transparent sphere?

\cite{Leite:2017zyb}

\sam{Sam to finish}

@@

\section{The Regge pole spectrum}

@@

 \subsection{The WKB approximation}

@@

\section{Scattering and CAM theory}
In this section we calculate the scattering cross section $d\sigma / d\Omega$ by means of the CAM method, and we compare with results obtained in the standard way from a partial-wave series \cite{Dolan:2017rtj}.

\subsection{The partial wave expansion}

@@

>>>>

Time-independent scattering by a compact body of radius $R$ with a regular centre, such as a neutron star or white dwarf, has received less attention, by comparison. In such a scenario, an electromagnetic wave will not penetrate far inside the compact body; but on the other hand, a gravitational wave will pass through the body without impediment from the matter distribution. Similarly, a neutron star is expected to be substantially transparent to neutrinos. Even in such cases where the coupling to matter is negligible, the incident wave will nevertheless be scattered by the influence of the spacetime curvature. Thus, the resulting scattering pattern depends not just on $M\omega$, but also on the internal structure of the body, and on its inverse compactness or \emph{tenuity}, $R/M$. Typical values of the tenuity are: $R/M \sim 6$ for neutron stars, $\sim 1.4-9.4 \times 10^3$ for white dwarfs, $4.7 \times 10^5$ for the Sun, and $1.4 \times 10^9$ for Earth. More speculatively, we shall also consider here the case of (hypothetical) \emph{ultra-compact objects} (UCOs), whose tenuity is bounded from below by the Buchdahl bound of $R / M = 9/4$.

Studies of black hole scattering typically involve the calculation of scattering amplitude(s) via partial-wave expansions.  A powerful approach for resumming these partial-wave expansions is provided by so-called Complex Angular Momentum (CAM) techniques. In the CAM approach, the sum over partial waves is replaced by an integral, and the contour of integration is deformed into  the complex-angular-momentum plane such that one collects a sum of residues of simple poles: these are the so-called \emph{Regge poles}.

@@

CAM techniques are yet to be fully applied to compact-body scattering, it appears, with one notable exception: in 1991, Chandrasekhar and Ferrari \cite{ChandrasekharIV:1992ey} examined the Regge pole spectra for relativistic stellar models. 
Their study focussed on resonant modes (poles with small imaginary parts corresponding to trapped $w$-modes or fluid modes), and it employed CAM theory to calculate the flow of gravitational energy through a star. It seems reasonable to assume that CAM theory can aid a deeper understanding of scattering by other compact objects, such as neutron stars and UCOs, and this is our purpose herein.

There is a close relationship between Regge poles and quasi-normal modes (QNMs); both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in \emph{time-dependent} scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a `ringdown phase' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well-studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

It was recently shown \cite{Dolan:2017rtj, Stratton:2019deq} that time-independent scattering by compact bodies generates (in principle) a \emph{rainbow scattering} phenomenon in the short wavelength regime ($M \omega \gg 1$). The rainbow scattering phenomenon is linked to the formation of a cusp caustic in the incident wavefront (see Fig.~1 in Ref.~\cite{Stratton:2019deq}), and to the stationary point in the geodesic deflection function associated with a ray (that is, a null geodesic) that passes somewhat inside the body. For less-dense bodies, such as stars and white dwarfs, the cusp caustic forms at a distance $d \sim R^2 / (4M)$ downstream of the body, and the rainbow angle is $\theta_r \sim 4M/R$ (e.g.~for the Sun, $d \approx 550 \, \text{a.u.}$ and $\theta_r \approx 1.8 \, \text{arcsec}$) \footnote{The rainbow angle $\theta_r \approx 4M/R$ is distinct from the Einstein ring angle of $\theta_E \approx \sqrt{4M/r}$}. For compact bodies such as neutron stars, however, the cusp caustic forms near the surface of the body, and the rainbow angle is large ($\theta_r \gtrsim 59.6^\circ$ for $R/M = 6$; see Table I in Ref.~\cite{Stratton:2019deq}). In this case, the rainbow angle is -- in principle at least -- a useful diagnostic of the matter distribution within the body and thus its nuclear equation of state.

@@

The remainder of this paper is organised as follows. Sec.~\ref{SecII} reviews the theory of scalar waves on a spherically-symmetric spacetime of a compact object. Here we describe our model spacetime (\ref{SecIIa}), the effective potentials for wave scattering (\ref{subsec:Potentials}) and the physical boundary conditions that define the S-matrix (\ref{subsec:bc}). Sec.~\ref{sec:RP} focusses on the Regge pole spectrum. Here we outline the link between Regge poles and QNMs (\ref{subsec:QNMs_RPs}), we describe the numerical method (\ref{subsec:method}), and we present new numerical results for the spectrum (\ref{subsec:results1}) which exhibits three distinct branches of Regge poles (see e.g.~Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}); in Sec.~\ref{subsec:WKB} we apply the WKB method to obtain an approximate formula for the `broad resonance' branch. Sec.~\ref{sec:CAM} concerns the application of CAM techniques to the calculation of the scattering cross section. Here we review the standard partial-wave expansion of the scattering amplitude (\ref{subsec:partial}), we apply CAM techniques to write this as the sum of a residue series and a background integral (\ref{SecIIc}), and we present a selection of numerical results for $d\sigma/d\Omega$ (\ref{subsec:results2}). We conclude with a discussion in Sec.~\ref{sec:conclusions}.

@@

\section{The Regge pole spectrum}\label{sec:RP}

@@

 \subsection{The WKB approximation}\label{subsec:WKB}

@@

\section{Scattering and CAM theory} \label{sec:CAM}
In this section we calculate the scattering cross section $d\sigma / d\Omega$ by means of the CAM method, and we compare with results obtained in the standard way from a partial-wave series \cite{Dolan:2017rtj}.

\subsection{The partial wave expansion} \label{subsec:partial}

@@

Finally a paragraph on the connections to Mie scattering by a transparent sphere.
